---
title: "Source apportionment Data analysis"
author: Erick E. Mollinedo
date: '`r format(Sys.Date(), "%B %d, %Y")`'
format: html
editor: visual
---

Packages needed for this script:

```{r}
library(readxl)
library(readr)
library(here)
library(tidyverse)
library(gt)
library(vtable)
library(rstatix)
library(ggpubr)
library(ggpattern)
library(treemap)
```

## INPUT DATA ANALYSES

### Description of samples

First, describing the number of filters per study arm and by visit used for source apportionment analysis

```{r}
#Load the HAPIN sample filters dataframe
hapin_filters <- read_csv(here("data", "processed-data", "hapin_samples.csv"))
```

Now, deleting the filter sample with missing PM2.5 and BC concentrations.

```{r}
#Delete observation with filter id: 3M53864.
hapin_filters <- hapin_filters %>% filter(filter_id != "3M53864")
```

Finally, creating a table that summarizes the number of filters per visit and study arm

```{r}
#Create data frame that summarizes the filter samples
summary_filters <- hapin_filters %>%
  count(arm, timepoint) %>%
  pivot_wider(names_from = timepoint, values_from = n, values_fill = list(n = 0))

#Rename arm to Study Arm
summary_filters <- summary_filters %>% rename("Study Arm" = arm)

#Save the table as .RDS file
saveRDS(summary_filters, file = here("output", "summary_filters.RDS"))

#Create table using gt
gt_filters <- summary_filters %>% gt() %>% 
  tab_spanner(label = "Time point", columns = c(BL, BLP1, P1, P1P2, P2, P2B1))

#View table
gt_filters
```

### Concentrations statistical summaries

First load the `concentration` data files.

```{r}
#Load the files
concentration <- read_csv(here("data", "processed-data", "concentration.csv")) #Only species for source app analysis
concentration_full <- read_csv(here("data", "processed-data", "concentrations_full.csv")) #All chemical species
```

Delete the missing concentration from the `concentration_full` data frame.

```{r}
#Delete observation with filter id: 3M53864.
concentration_full <- concentration_full %>% filter(filter_id != "3M53864")
```

Create a summary statistics table for all the chemical species, by study arm, and including baseline measurements sepparately. This is done using the `vtable` package. After, a summary statistics table with the chemical species selected, first overall and then by study group.

```{r}
#ALL CHEMICAL SPECIES
#Create a summary table, this divides into BL, Intervention and Control
sumtable(concentration_full, group = 'group', group.long = T, add.median = T)

#Create a summary table, but with all observations overall
sumtable(concentration_full, add.median = T)

#FILTERED CHEMICAL SPECIES
#Create a summary table, this divides into BL, Intervention and Control
sumtable(concentration, group = 'group', group.long = T, add.median = T)

#Create a summary table, but with all observations overall
sumtable(concentration, add.median = T)
```

Make some edits to the `concentration` dataframe to run the statistical analyses.

```{r}
#Exclude the arm and fueltype variables
concentration_filtered <- concentration %>% select(-c(arm, fueltype))

#Pivot longer for boxplots
options(scipen = 999) #to remove scientific notation for smaller values (i.e. changes from 1*10-3 to 0.001)
concentration_longer <- concentration_filtered %>% 
  pivot_longer(cols = -c(filter_id, group),
               names_to = "species", values_to = "concentration")

#Set order of species for factor conversion
#level_order <- c("Mg", "Al", "Si", "S", "K", "Ca", "Ti", "Mn", "Fe", "Zn", "Cl", "Na", "BC", "PM25")
level_order <- c("Zn", "Mn", "Ti", "Na", "Mg", "Fe", "S", "Ca", "Al", "Si", "K", "Cl", "BC", "PM25")

#Convert species to factors and assign order specified before
concentration_longer$species <- factor(concentration_longer$species, levels = level_order)
```

Compute the paired Wilcoxon test. First, creating a function.

```{r}
#Function to perform the Wilcoxon test
perform_wilcoxon <- function(species_data) {
  wilcox.test(concentration ~ group, data = species_data, subset = group %in% c("Intervention", "Control"))
}

#Apply the function to each species
results_wilcoxon <- by(concentration_longer, concentration_longer$species, perform_wilcoxon)

print(results_wilcoxon)
```

Create boxplots to visualize the differences in concentrations for each chemical species. First, doing the t-test to assess differences between study arm.

```{r}
#Conduct the t-test of concentration-fuel_type (Use the rstatix package)
stat.test <- concentration_longer %>% 
  filter(group != "Baseline") %>% #To create the statistical significance only between Control and Intervention
  group_by(species) %>% 
  t_test(concentration ~ group) %>% 
  add_significance(cutpoints = c(0, 0.05, 1), symbols = c("**", "ns")) #to add the significance values

#Add the position coordinates
stat.test <- stat.test %>% 
  add_x_position(x= "species", dodge = 0.5) %>% #x position is species, and the length of significance bar is 0.5
  add_y_position(y.trans = function(x){log10(x)}) #add y position, in this case specify using a function that it is in log base 10 scale
```

Now, creating a plot in black and white, this is the one used for the manuscript.

```{r}
#Create a pattern
studyarm_patterns <- c("Baseline" = "none",
                       "Control" = "stripe",
                       "Intervention" = "circle")

#Boxplot with patterns instead of colors
concentration_longer %>%
  ggplot() +
  geom_boxplot_pattern(aes(x = species, y = concentration, pattern = group),
    fill = "white",
    colour = "black",
    pattern_fill = "black",
    pattern_density = 0.1,
    pattern_spacing = 0.01) +
  scale_pattern_manual(values = studyarm_patterns) +
  xlab(expression(Chemical~species)) +
  ylab(expression(Concentration~(ug/m^3))) +
  scale_y_log10(breaks = c(0.01, 0.1, 1.0, 10.0, 100.0, 1000.0), 
                labels = c(0.01, 0.1, 1, 10, 100, 1000))+
  scale_x_discrete(labels = c("Zn", "Mn", "Ti", "Na", "Mg", "Fe", "S", "Ca",
                              "Al", "Si", "K", "Cl", "BC", "PM2.5")) +
  guides(pattern = guide_legend(title = "Study Group")) +
  theme_test(base_size = 14) +
  stat_pvalue_manual(stat.test, label = "p.signif", tip.length = 0.004)

#Save plot using ggsave
ggsave(here("figures", "figure1_b&w.png"), plot = last_plot(), width = 20, height = 10, units = "cm")
```

If the Figure above needs to be in color, use the code below:

```{r}
#Create color palette for the boxplots
studyarm_colors <- c("Baseline" = "skyblue",
                   "Control" = "orangered2",
                   "Intervention" = "plum3")

#Create the boxplot (in color)
concentration_longer %>% #filter(group != "BL") %>% (In case just want to show Control-Intervention post-randomization)
ggplot()+
  geom_boxplot(aes(x= species, y= concentration, fill= group))+
  scale_fill_manual(values = studyarm_colors)+ #Apply the color palette created above
  xlab(expression(Chemical~species))+ #Label the x-axis
  ylab(expression(Concentration~(ug/m^3)))+ #Label the y-axis
  scale_y_log10(breaks= c(0.01, 0.1, 1.0, 10.0, 100.0, 1000.0), labels= c(0.01, 0.1, 1, 10, 100, 1000))+
  scale_x_discrete(labels= c("Mg", "Al", "Si", "S", "K", "Ca", "Ti", "Mn", "Fe", "Zn", "Cl", "Na", "BC", "PM2.5"))+
  scale_x_discrete(labels= c("Zn", "Mn", "Ti", "Na", "Mg", "Fe", "S", "Ca", "Al", "Si", "K", "Cl", "BC", "PM2.5"))+
  guides(fill= guide_legend(title= "Study Group"))+ #Change legend from group to Study arm
  theme_bw(base_size = 14)+ #Select the theme and size of x and y labels
  stat_pvalue_manual(stat.test, label = "p.signif", tip.length = 0.004) #Add the significance bars, use 'p.signif' to add *** or 'ns'

#Save plot using ggsave
ggsave(here("figures", "figure1.png"), plot = last_plot(), width = 20, height = 10, units = "cm")
```


## OUTPUT DATA ANALYSES

### Q(True)/Q(Exp) plot

Now I will create a plot that explains the Qtrue/Qexp ratio according to the number of factors. First, I will load the `q-factors` csv file that comes from the diagnostics files.

```{r}
#Loading the data for Q plot
q_data <- read_csv(here("data", "raw-data", "q-factors.csv"))

#Exploring the df
str(q_data)

#Make the Factors as `factors`
as.factor(q_data$fact)
```

Now, creating a plot using geom_point and geom_line.

```{r}
#Create a Qtrue/Qexp ratio plot
ggplot(q_data, aes(x= fact, y= ratio))+
  geom_point(size= 4, color= "firebrick")+
  geom_line()+
  theme_bw()+
  labs(x= "Number of Factors", y= "Q(true)/Q(exp)")+
  ggtitle("Q(True)/Q(Exp) plot by Number of factors")+
  theme(axis.text = element_text(size = 12),
        axis.title = element_text(size = 14),
        plot.title = element_text(size= 17, hjust = 0.5))+
  ylim(1, 18)

#Save plot using ggsave
ggsave(here("figures", "qqplot.png"), plot = last_plot(), width = 20, height = 12, units = "cm")
```

### Factor profiles

The following part is to create a plot that is visually more clear than the plot obtained from the EPA PMF 5.0 software. First, load the `species` data, which is obtained from the '% of species sum' section from the `HAPIN4_base` output file of the software, and then perform some edits.

```{r}
#Load the species concentrations df, obtained from the PMF analysis
species <- read_csv(here("data", "raw-data", "species.csv"))

#Pivot longer
spec_longer <- species %>% 
  pivot_longer(cols = -species,
               names_to = "Factor", values_to = "Percent")

#Rename the columns
spec_longer <- spec_longer %>% 
  mutate(Factor = case_when(Factor == "factor_1" ~ "Factor 1",
                            Factor == "factor_2" ~ "Factor 2",
                            Factor == "factor_3" ~ "Factor 3",
                            Factor == "factor_4" ~ "Factor 4"))
```

First, set the order of the species, and set as factor object.

```{r}
#Set new level order
level_order2 <- c("Mg", "Al", "Si", "S", "K", "Ca", "Ti", "Mn", "Fe", "Zn", "Cl", "Na", "BC", "PM2.5")

#Convert species to factors and assign order specified before
spec_longer$species <- factor(spec_longer$species, levels = level_order2)
```

Now, create a graph of profiles (In black and white) by each of the four factors.

```{r}
#Create the Black and white gradient
factor_colors <- c("Factor 1" = "grey75",
                   "Factor 2" = "grey50",
                   "Factor 3" = "grey25",
                   "Factor 4" = "grey0")

#Create the plot
ggplot(spec_longer, aes(x= species, y= Percent, fill= Factor))+
  geom_col()+
  facet_grid(Factor ~., scales = "fixed")+ #Set the grid and make the scale the same for all Factors
  scale_y_continuous(position = "left")+ #Set the y-scale axis on the left
  labs(x= "Species", y= "Proportion of Species (%)")+ #Change the labels
  scale_x_discrete(labels= c("Mg", "Al", "Si", "S", "K", "Ca", "Ti", "Mn", "Fe", "Cl", "Na", "BC", "PM2.5"))+ #Manually label the x-axis
  #ggtitle("Species Profiles by a Four-factor model")+
  scale_fill_manual(values = factor_colors, #Set the color palette created before
                    labels= c("Factor 1" = "Biomass", #Rename factors to their corresponding source names
                              "Factor 2" = "Crustal",
                              "Factor 3" = "Other soil",
                              "Factor 4" = "Fossil fuel"))+
  guides(fill= guide_legend(title= "Source"))+ #Change legend from Factor to Source
  theme_test()+
  theme(axis.text = element_text(size = 11),
        axis.title = element_text(size = 14),
        plot.title = element_text(size = 18, hjust = 0.5),
        legend.title = element_text(hjust = 0.5), #adjust legend title (center)
        strip.text.y = element_blank(),) #remove the grey box labels for each factor profile

#Save plot using ggsave
ggsave(here("figures", "figure2_b&w.png"), plot = last_plot(), width = 8, height = 6, units = "in", dpi = 1000)
```

If the graph is needed in colors, use the code below:

```{r}
# Define color palette for factors
factor_colors <- c("Factor 1" = "cyan4",
                   "Factor 2" = "indianred2",
                   "Factor 3" = "darksalmon",
                   "Factor 4" = "lightskyblue")

#Create the plot
ggplot(spec_longer, aes(x= species, y= Percent, fill= Factor))+
  geom_col()+
  facet_grid(Factor ~., scales = "fixed")+ #Set the grid and make the scale the same for all Factors
  scale_y_continuous(position = "left")+ #Set the y-scale axis on the left
  labs(x= "Species", y= "Proportion of Species (%)")+ #Change the labels
  scale_x_discrete(labels= c("Mg", "Al", "Si", "S", "K", "Ca", "Ti", "Mn", "Fe", "Cl", "Na", "BC", "PM2.5"))+ #Manually label the x-axis
  #ggtitle("Species Profiles by a Four-factor model")+
  scale_fill_manual(values = factor_colors, #Set the color palette created before
                    labels= c("Factor 1" = "Biomass", #Rename factors to their corresponding source names
                              "Factor 2" = "Crustal",
                              "Factor 3" = "Other soil",
                              "Factor 4" = "Fossil fuel"))+
  guides(fill= guide_legend(title= "Source"))+ #Change legend from Factor to Source
  theme_test()+
  theme(axis.text = element_text(size = 11),
        axis.title = element_text(size = 14),
        plot.title = element_text(size = 18, hjust = 0.5),
        legend.title = element_text(hjust = 0.5), #adjust legend title (center)
        strip.text.y = element_blank(),) #remove the grey box labels for each factor profile

#Save plot using ggsave
ggsave(here("figures", "figure2.png"), plot = last_plot(), width = 8, height = 6, units = "in", dpi = 1000)
```


### PM2.5 contributions

Now, to check the proportion of PM2.5 that is contributed by each study arm (including baseline), and divided by source. For this part, I will work with the `concentration` data frame. First I will do some edits, so I keep only the observed PM2.5 concentrations and other variables. Then I will estimate the concentrations apportioned to each source (factor), by using the values from the `HAPIN4_base` output from the EPA PMF 5.0 software. Finally, the 95% CIs from the concentrations will be reported.

```{r}
#Select only the ID's, observed PM2.5 concentrations, fueltype, arm and group variables
pm25_contributions <- concentration %>% 
  select(c(filter_id, arm, fueltype, group, PM25))

#Rename 'PM25' to 'obs_pm'
pm25_contributions <- pm25_contributions %>% rename(obs_pm = "PM25")

#Calculate the mass of PM2.5 for each source (factor)
pm25_contributions <- pm25_contributions %>% 
  mutate(source_1 = obs_pm * 0.6964239, #69.64% of PM2.5 contributes to factor 1
         source_2 = obs_pm * 0.04120215, #4.12% of PM2.5 contributes to factor 2
         source_3 = obs_pm * 0.04211482, #4.21% of PM2.5 contributes to factor 3
         source_4 = obs_pm * 0.2202591) #22.02% of PM2.5 contributes to factor 4

#ESTIMATE CONTRIBUTIONS FOR CONTROL, INTERVENTION AND BASELINE ONLY
# Calculate the sum of the variables from 'source_1' to 'source_4'
contributions_group <- pm25_contributions %>%
  mutate(sum_pm = rowSums(select(., source_1:source_4), na.rm = TRUE))

# Function to calculate mean, standard error, and 95% confidence interval
calculate_stats <- function(data) {
  mean_val <- mean(data, na.rm = TRUE)
  se_val <- sd(data, na.rm = TRUE) / sqrt(length(data))
  ci_lower <- mean_val - qt(0.975, df = length(data) - 1) * se_val
  ci_upper <- mean_val + qt(0.975, df = length(data) - 1) * se_val
  return(data.frame(mean = mean_val, se = se_val, ci_lower = ci_lower, ci_upper = ci_upper))
}

# Group by 'group' and calculate the required statistics, including 95% CIs
summary_group <- contributions_group %>%
  group_by(group) %>%
  summarise(across(sum_pm, list(mean = ~mean(.), se = ~sd(.) / sqrt(n()), 
                                ci_lower = ~mean(.) - qt(0.975, df = n() - 1) * (sd(.) / sqrt(n())), 
                                ci_upper = ~mean(.) + qt(0.975, df = n() - 1) * (sd(.) / sqrt(n())))))

# Display the results
print(summary_group)


#FOR SOURCE 1, 2, 3 4 BY STUDY GROUP
# Compute mean, standard error, and 95% confidence intervals of the PM2.5 concentrations grouped by source and by study group
#summary_sources <- pm25_contributions %>%
  #group_by(group) %>% #Group by study group
  #summarise(
    #mean_source_1 = mean(source_1, na.rm = TRUE), #Estimate the mean for source 1
    #se_source_1 = sd(source_1, na.rm = TRUE) / sqrt(n()), #Estimate the SD for source 1
    #ci_lower_source_1 = mean_source_1 - qt(0.975, df = n() - 1) * se_source_1, #Estimate the lower limit of source 1
    #ci_upper_source_1 = mean_source_1 + qt(0.975, df = n() - 1) * se_source_1, #Estimate the upper limit of source 2
    
    #mean_source_2 = mean(source_2, na.rm = TRUE),
    #se_source_2 = sd(source_2, na.rm = TRUE) / sqrt(n()),
    #ci_lower_source_2 = mean_source_2 - qt(0.975, df = n() - 1) * se_source_2,
    #ci_upper_source_2 = mean_source_2 + qt(0.975, df = n() - 1) * se_source_2,
    
    #mean_source_3 = mean(source_3, na.rm = TRUE),
    #se_source_3 = sd(source_3, na.rm = TRUE) / sqrt(n()),
    #ci_lower_source_3 = mean_source_3 - qt(0.975, df = n() - 1) * se_source_3,
    #ci_upper_source_3 = mean_source_3 + qt(0.975, df = n() - 1) * se_source_3,
    
    #mean_source_4 = mean(source_4, na.rm = TRUE),
    #se_source_4 = sd(source_4, na.rm = TRUE) / sqrt(n()),
    #ci_lower_source_4 = mean_source_4 - qt(0.975, df = n() - 1) * se_source_4,
    #ci_upper_source_4 = mean_source_4 + qt(0.975, df = n() - 1) * se_source_4)

# Print the results
#print(summary_sources)

#FOR SOURCE 1, 2, 3 4 ONLY
# Transform the data set (PART 2)
#summary_transformed <- pm25_contributions %>%
  #pivot_longer(cols = starts_with("source_"), names_to = "Source", values_to = "Value") %>%
  #pivot_wider(names_from = group, values_from = Value)

#Compute mean, standard error, and 95% confidence intervals of the PM2.5 concentrations grouped by study group
#summary_group <- summary_transformed %>%
  #group_by(Source) %>% #Group by Source
  #summarise(
    #mean_baseline = mean(Baseline, na.rm = TRUE), #Estimate the mean for baseline
    #se_baseline = sd(Baseline, na.rm = TRUE) / sqrt(n()), #Estimate the SD for baseline
    #ci_lower_baseline = mean_baseline - qt(0.975, df = n() - 1) * se_baseline, #Estimate the lower limit of baseline
    #ci_upper_baseline = mean_baseline + qt(0.975, df = n() - 1) * se_baseline, #Estimate the upper limit of baseline
    
    #mean_control = mean(Control, na.rm = TRUE),
    #se_control = sd(Control, na.rm = TRUE) / sqrt(n()),
    #ci_lower_control = mean_control - qt(0.975, df = n() - 1) * se_control,
    #ci_upper_control = mean_control + qt(0.975, df = n() - 1) * se_control,
    
    #mean_intervention = mean(Intervention, na.rm = TRUE),
    #se_intervention = sd(Intervention, na.rm = TRUE) / sqrt(n()),
    #ci_lower_intervention = mean_intervention - qt(0.975, df = n() - 1) * se_intervention,
    #ci_upper_intervention = mean_intervention + qt(0.975, df = n() - 1) * se_intervention)

# Print the results
#print(summary_group)
```

#### PM2.5 contributions graphs

Do bar charts that show the contribution of each source by each group (BL, Control, Intervention). First, create a data frame with the mean contributions of PM2.5 by source and group.

```{r}

# Calculate the mean of the factor variables by 'arm'
mean_contributions <- pm25_contributions %>%
  group_by(group) %>%
  summarise(mean_source1 = mean(source_1, na.rm = TRUE),
            mean_source2 = mean(source_2, na.rm = TRUE),
            mean_source3 = mean(source_3, na.rm = TRUE),
            mean_source4 = mean(source_4, na.rm = TRUE))

#Reshape contributions data frame
contributions_summary <- mean_contributions %>% 
  pivot_longer(cols = c(mean_source1, mean_source2, mean_source3, mean_source4),
               names_to = "source",
               values_to = "concentration")

#Estimate total percentage contributions by group and factor
contributions_summary <- contributions_summary %>% 
  mutate(percentage = concentration/sum(concentration) * 100) %>% 
  mutate(source = case_when(
    source == "mean_source1" ~ "source1",
    source == "mean_source2" ~ "source2",
    source == "mean_source3" ~ "source3",
    source == "mean_source4" ~ "source4",
    TRUE ~ source))

#Calculate the standard errors by group
#contributions_summary <- contributions_summary %>%
  #group_by(group) %>%
  #mutate(error = sd(concentration) / sqrt(n()))

#Add confidence intervals to data frame
contributions_summary$ci_lower <- c(rep(129.74, 4), rep(114.71, 4), rep(27.18, 4))
contributions_summary$ci_upper <- c(rep(180.15, 4), rep(139.67, 4), rep(32.66, 4))

#Print dataframe
print(contributions_summary)
```

Create a plot in black and white.

```{r}
#Define the patter for each factor
factor_patterns <- c("source1" = "circle",
                       "source2" = "stripe",
                       "source3" = "none",
                     "source4" = "crosshatch")

#Create plot (with patterns instead of grayscale)
ggplot(contributions_summary, aes(x = group, y = concentration, pattern = source)) +
  geom_col_pattern(fill = "white", color = "black", pattern_fill = "black", #keep bars white and pattern color
                   pattern_colour = "black", pattern_density = 0.2, pattern_spacing = 0.02) + 
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), 
                width = 0.2, color = "black") +  # Add error bars
  scale_y_continuous(breaks = seq(0, 220, by = 10)) + #Scale of the y-axis
  labs(y = expression(Mass~concentration~(ug/m^3)), x = "Study group") + #Relabel x and y axis
  scale_pattern_manual(values = factor_patterns,
                    labels= c("source1" = "Biomass",
                              "source2" = "Crustal",
                              "source3" = "Other soil",
                              "source4" = "Fossil fuel"))+
  theme_test()+
  theme(axis.text = element_text(size = 9), #Change the size of the x and y scale and labels
        axis.title = element_text(size = 14),
        axis.text.x = element_text(size = 12),
        legend.title = element_text(hjust = 0.5),
        legend.text = element_text(size = 10))

#Save plot using ggsave
ggsave(here("figures", "figure5_b&w.png"), plot = last_plot(), width = 20, height = 14, units = "cm")
```

Same plot from above, but in colors:

```{r}
#Define factor colors
factor_colors2 <- c("source1" = "cyan4",
                   "source2" = "indianred2",
                   "source3" = "darksalmon",
                   "source4" = "lightskyblue")

#Create plot (on a grayscale)
ggplot(contributions_summary, aes(x = group, y = concentration, fill = source)) +
  geom_col(color= NA) + #Color the borders of the bars
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), 
                width = 0.2, color = "black") +  # Add error bars
  scale_y_continuous(breaks = seq(0, 220, by = 10)) + #Scale of the y-axis
  labs(y = expression(Mass~concentration~(ug/m^3)), x = "Study group") + #Relabel x and y axis
  scale_fill_manual(values = factor_colors2,
                    labels= c("source1" = "Biomass",
                              "source2" = "Crustal",
                              "source3" = "Other soil",
                              "source4" = "Fossil fuel"))+ #Input the color palette manually
  theme_test()+
  theme(axis.text = element_text(size = 9), #Change the size of the x and y scale and labels
        axis.title = element_text(size = 14),
        axis.text.x = element_text(size = 12),
        legend.title = element_text(hjust = 0.5),
        legend.text = element_text(size = 10))

#Save plot using ggsave
ggsave(here("figures", "figure5.png"), plot = last_plot(), width = 20, height = 14, units = "cm")
```


Do something similar as a pie chart, but improved, to visualize the contribution of PM2.5 concentrations by study arm (including baseline) First building the dataset for the treemap by sources.

```{r}
#Build data set
group <- c("Biomass", "Crustal", "Other soil", "Fossil fuels") #Sources as groups
value <- c(54.25, 3.21, 3.28, 17.16) #Input concentrations as values, this comes from the HAPIN4_base xlsx file. List them in order
treedata <- data.frame(group, value) #Create the treemap data frame
```

Now plotting the treemap of study arm by sources. The graph from below will be manually saved, so then it will be edited in paint to create the graphical abstract.

```{r}
#Define palette of colors
treemap_colors <- c("#00FF99", "#FF6666", "#66CCFF", "#FFCC99")

#Create the plot (Note: The concentrations and % values will be entered manually on paint)
treemap(treedata, #Select the dataset
        index = "group", #Select the group and subgroup values to plot the quadrants
        vSize = "value", #Select the values that give the sizes for each group and subgroup
        type = "index", #The colors are determined by the index variables
        title = expression(PM[2.5]~mass~contribution~to~each~source), #Add title
        fontsize.labels = 0, #size of the group labels, 0, so it manually is edited later
        fontcolor.labels = "black",
        fontsize.title = 18, #Size for the title
        fontsize.legend = 10, #Not sure if this is necessary
        bg.labels = "transparent", #Only show the labels with no highlights around the labels
        palette = treemap_colors, #Select the color palette manually
        align.labels = list(c("center", "center"), c("center", "bottom"))) #Placement for the labels, first is group and second is subgroups

```

### Pie chart of PM2.5 contributions by source

The first pie chart is in black and white:

```{r}
#Load the dataset
pm25 <- read_csv(here("data", "raw-data", "pm25_proportions.csv"))

#Define the color palette (Black and white)
factor_colors3 <- c("Biomass" = "grey75",
                   "Crustal" = "grey50",
                   "Other Soil" = "grey25",
                   "Fossil Fuel" = "grey0")

#Order for sources
pm25$Source <- factor(pm25$Source,
                      levels = c("Biomass", "Crustal", "Other Soil", "Fossil Fuel"))

#Compute the position of the labels
pm25 <- pm25 %>% 
  arrange(desc(Source)) %>%
  mutate(prop = round(Concentration / sum(pm25$Concentration) *100, 2)) %>%
  mutate(ypos = cumsum(prop)- 0.6*prop )

# Graph a piechart
ggplot(pm25, aes(x="", y= prop, fill= Source)) +
  geom_bar(stat="identity", width=1, color= "white") +
  coord_polar("y", start=0)+
  scale_fill_manual(values = factor_colors3)+
  theme_void()+
  theme(legend.position = "left", 
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14, hjust = 0.5),
        plot.title = element_text(size = 16, hjust = 0.5))

#Then, manually enter the concentrations and proportions in paint

#Save plot using ggsave
ggsave(here("figures", "figure3_b&w.png"), plot = last_plot(), width = 20, height = 14, units = "cm", bg = "white")

```

This code below for the pie chart in colors:

```{r}
#Define the color palette
factor_colors3 <- c("Biomass" = "cyan4",
                   "Crustal" = "indianred2",
                   "Other Soil" = "darksalmon",
                   "Fossil Fuel" = "lightskyblue")

#Order for sources
pm25$Source <- factor(pm25$Source,
                      levels = c("Biomass", "Crustal", "Other Soil", "Fossil Fuel"))

#Compute the position of the labels
pm25 <- pm25 %>% 
  arrange(desc(Source)) %>%
  mutate(prop = round(Concentration / sum(pm25$Concentration) *100, 2)) %>%
  mutate(ypos = cumsum(prop)- 0.6*prop )

# Graph a piechart
ggplot(pm25, aes(x="", y= prop, fill= Source)) +
  geom_bar(stat="identity", width=1, color= "white") +
  coord_polar("y", start=0)+
  scale_fill_manual(values = factor_colors3)+
  theme_void()+
  theme(legend.position = "left", 
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14, hjust = 0.5),
        plot.title = element_text(size = 16, hjust = 0.5))

#Then, manually enter the concentrations and proportions in paint

#Save plot using ggsave
ggsave(here("figures", "figure3.png"), plot = last_plot(), width = 20, height = 14, units = "cm", bg = "white")
```


### Observed-Predicted PM2.5

This part is computed using the source contributions of PM2.5 for each factor (source). In summary, a .csv file that contains the filter IDs and the PM2.5 contributions from each factor was created by copying the values from row 7590 from the `HAPIN4_sourcecontributions.xls` output file from the EPA PMF 5.0 software. Then, the total PM2.5 was estimated as the sum of PM2.5 contributions by each factor, and the observed PM2.5 from the `concentrations` data file were also placed into the file.

```{r}
#Load data
predpm_contributions <- read_csv(here("data", "processed-data", "contributions.csv"))

#Sum the contributions from each factor and add the observed PM2.5 concentrations
predpm_contributions <- predpm_contributions %>% 
  mutate(predpm = rowSums(select(., factor_1:factor_4))) %>% #Sum contributions
  mutate(obspm = concentration$PM25) #Add observed PM2.5 values
```

Now, do a graph that plots the observed vs predicted PM2.5 mass concentrations

```{r}
#Do a graph of obs vs pred values using ggplot()
predpm_contributions %>% 
  ggplot(aes(x= obspm, y= predpm))+ #Selecting the observed and predicted PM2.5 concentrations
  geom_point(shape= 21, color= "black", size= 2)+ #Select point as geometry and size and color of the point
  geom_smooth(method = lm, colour= "black", se= T)+ #Add the lm line with SE
  xlab(expression(Observed~PM[2.5]~concentration~(ug/m^3)))+ #X label
  ylab(expression(Predicted~PM[2.5]~concentration~(ug/m^3)))+ #y label
  scale_x_continuous(limits = c(0, 800))+ #Manually set the x-limits
  scale_y_continuous(limits = c(0, 800))+ #Manually set the y-limits
  #labs(title = expression(Observed~vs~Predicted~PM[2.5]~mass~concentrations))+ #Title of the plot
  theme_test()+
  theme(aspect.ratio = 1)+ #Aspect ratio of graph "square"
  geom_abline(intercept = 0, slope = 1, linewidth= 0.5, linetype = "dashed")+ #set an abline dashed
  stat_cor(aes(label = paste(after_stat(rr.label), sep = "~`, `~")), label.x = 3) + # Add R2 value
  #stat_cor(method = "pearson", #add the correlation
           #alternative = "two.sided", #two-sided test
           #cor.coef.name = expression(R^2), #for Spearman symbol
           #label.x.npc = "left", #x position
           #label.y.npc = "top", #y position
           #p.accuracy = 0.001, #accuracy of the p-value
           #size= 4)+ #size of text
  annotate("text", x= 230, y= 700, label= paste("y= (0.62 ± 0.02)x + (20.78 ± 2.12)"), size= 4)

#Save plot
ggsave(here("figures", "figure5.png"), plot = last_plot(), width = 20, height = 14, units = "cm")
```

And now do a time-series like plot for the observed vs predicted PM2.5 concentrations

```{r}
#Reshape data frame with pivot_longer()
pm_long <- predpm_contributions %>%
  pivot_longer(cols = c(predpm, obspm), names_to = "type", values_to = "value")

#Set color palette for the plot
pm_palette <- c("#E66100", "#5D3A9B")
pm_labels <- c("obspm"= "Observed", "predpm"= "Predicted")

#Create plot
ggplot(pm_long, aes(x = filter_id, y = value, color = type, group = type)) +
  geom_point() +
  geom_line() +
  labs(title = expression(Observed~vs~Predicted~PM[2.5]~mass~concentrations),
       color = "Type") +
  xlab(expression(Data~Point))+ #X label
  ylab(expression(PM[2.5]~mass~concentration~(ug/m^3)))+ #y label
  scale_color_manual(values= pm_palette, labels= pm_labels)+
  theme_minimal()+
  theme(axis.text.x = element_blank(),
        plot.title = element_text(hjust = 0.5))

ggsave(here("figures", "obs-pred.png"), plot = last_plot(), width = 20, height = 14, units = "cm")
```

